---
title: "Pre-Process"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pre_Process}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

```{r setup}
#library(idepGolem)
devtools::load_all()
```

```{r message=FALSE, warning=FALSE, include=FALSE}
# Make data for pre-processing functions
idep_data <- get_idep_data()

YOUR_DATA_PATH <- "E:/idep_9_24/data/data103/data_go/BcellGSE71176_p53.csv"
YOUR_EXPERIMENT_PATH <- "E:/idep_9_24/data/data103/data_go/BcellGSE71176_p53_sampleInfo.csv"

expression_file <- data.frame(
  datapath = YOUR_DATA_PATH
)
experiment_file <- data.frame(
  datapath = YOUR_EXPERIMENT_PATH
)

load_data <- input_data(
  expression_file = expression_file,
  experiment_file = experiment_file,
  go_button = FALSE,
  demo_data_file = idep_data$demo_data_file,
  demo_metadata_file = idep_data$demo_metadata_file
)

converted <- convert_id(
  query = rownames(load_data$data),
  idep_data = idep_data,
  select_org = "BestMatch"
)

all_gene_info <- gene_info(
  converted = converted,
  select_org = "BestMatch",
  idep_data = idep_data
)

converted_data <- convert_data(
  converted = converted,
  no_id_conversion = FALSE,
  data = load_data$data
)

gene_names <- get_all_gene_names(
  mapped_ids = converted_data$mapped_ids,
  all_gene_info = all_gene_info
)
```

This instruction builds off the objects and functions from the "Load_Data"
vignette. Please see that document before using the functions described below.
We will now transform and process the data based off the type that is being
studied. The package also provides visualization functions that will provide a
foundation for exploratory data analysis and good overview of the data.

#### **`pre_process`** Function

Processing the data requires many options to ensure the data gets to the desired
form. This is the data that will be used in the exploratory analysis to come.
The data to be processed is the `converted_data` from above. The parameters
`missing_value` and `data_file_format` have a list of input values to select 
from for processing. These values and descriptions can be found in the tables 
below. `missing_value` determines how missing data entries will be filled in, 
and `data_file_format` tells the function what type of data is being processed.

<div style="text-align: left">

Input (`missing_value`)    Description
-------------------------  -----------------------------------------------------------
`geneMedian`               Substitute median expression value for gene across all samples
`treatAsZero`              Substitute "0" for all missing values
`geneMedianInGroup`        Substitute median expression value for gene within sample group

Input (`data_file_format`)   Description
---------------------------  -----------------------------------------------------------------------
`1`                          Read counts data (recommended)
`2`                          Normalized expression values (RNA-seq FPKM, micro-array, etc.)
`3`                          Fold-changes and corrected P values from CuffDiff or any other program

</div>

<br>

##### **Read counts data**

For read counts data, the parameters `min_counts`, `n_min_samples_counts`,
`counts_transform` and `counts_log_start` need input values. `min_counts` and
`n_min_samples` work to filter the genes out of the data that do not have
sufficient counts. The function counts the columns that have more counts than
`min_counts` for each gene and filters out the genes with less columns than
`n_min_samples`. For example, if you only wanted genes with more than 0 counts
for each sample, you would set `min_counts = .5` and 
`n_min_samples = ncol(converted_data)`. `counts_transform` specifies the
transformation to perform on the counts data. This parameter has 3 input options
which are outlined below. If `counts_transform = 1`, `counts_log_start`
determines the constant to add to CPM to ensure that there are no `log(0)`
errors. All other parameters can be set to `NULL`

Input (`counts_transform`)   Description
---------------------------  --------------------------------------------------
`1`                          EdgeR: log2(CPM+c) (Counts per Million + constant) (Visit https://www.rdocumentation.org/packages/SparkR/versions/2.1.2/topics/log2)
`2`                          VST: variance stabilizing transform (Visit https://www.rdocumentation.org/packages/DESeq2/versions/1.12.3/topics/vst)
`3`                          rlog: regularized log (slow) (Visit https://www.rdocumentation.org/packages/DESeq2/versions/1.12.3/topics/rlog)

<br>

##### **Normalized expression values Parameters**

If the data being processed is for normalized expression values, the parameters
`low_filter_fpkm`, `n_min_samples_fpkm`, `log_transform_fpkm` and 
`log_start_fpkm` need input values. `low_filter_fpkm` and `n_min_samples_fpkm`
work together to filter out the genes that do not have enough counts. The
function will count the columns that are greater than `low_filter_fpkm` and 
discard the gene if it is not more than `n_min_samples_fpkm` columns. The
parameter `log_transform_fpkm` is `TRUE/FALSE` for whether to perform a log
transform on the data or not. If `log_transform = TRUE`, then `log_start_fpkm`
is the value that will be added to ensure there are no `log(0)` errors. All
other values can be set to `NULL`.

<br>

##### **Fold-changes and corrected P values**

For this data type, there is only one additional parameter to provide an input
for, `no_fdr`. This is a logical `TRUE/FALSE` to specify whether the data
contains corrected p-values or not. `TRUE` denotes fold-change only data. All
other parameters can be set to `NULL`.

<br>
<br>

The demonstration loaded in the first instruction is read counts data. The code
below shows an example use of the `pre_process` function using the example data.
Again, the data that is used is the data with the converted ensembl IDs. We will
return a large list of the results of the processing from this function. The
objects in this list can be found in the table below.

```{r}
processed_data <- pre_process(
  data = converted_data$data,
  missing_value = "geneMedian",
  data_file_format = 1,
  low_filter_fpkm = NULL,
  n_min_samples_fpkm = NULL,
  log_transform_fpkm = NULL,
  log_start_fpkm = NULL,
  min_counts = .5,
  n_min_samples_count = 1,
  counts_transform = 1,
  counts_log_start = 4,
  no_fdr = NULL
)
```

