---
title: "Clustering Workflow"
output: html_document
params:
  pre_processed_data:
  all_gene_names:
  sample_info:
  n_genes:
  k_clusters:
  cluster_meth:
  select_gene_id:
  list_factors_heatmap:
  heatmap_color_select:
  dist_function:
  hclust_function:
  heatmap_cutoff:
  gene_centering:
  gene_normalize:
  no_sample_clustering:
  show_row_dend:
  printcode:
    label: "Display Code"
    value: TRUE
    input: checkbox
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = params$printcode)
```

<!-- This document cannot be knit directly from R studio -->

This html document contains what parameters values were selected on the IDEP interface of the "Clustering" tab. It also includes the plots generated from those selections.

# ALL SELECTED PARAMETERS

```{r, include=TRUE}
for (i in 1:length(params)){
  # exclude loaded data & sample info
  if(names(params)[i] != "sample_info" && names(params)[i] != "pre_processed_data" && names(params)[i] != "all_gene_names")
    cat(paste0(names(params)[i], ": ", params[[i]], "\n") )
}
```

# Citation

The following plots were generated using IDEP with the following filters:
Include only the top `r params$n_genes` genes
using `r params$select_gene_id` to label
color scheme: `r params$heatmap_color_select `
distance: `r switch(params$dist_function,
"1" = "Hierarchical",
"2" = "K-Means")`
Linkage: `r params$hclust_function`
Cut off Z-score: `r params$heatmap_cutoff`
Center Genes: `r params$gene_centering`
Gene normalize: `r params$gene_normalize`
Do not cluster samples: `r params$no_sample_clustering`
Show row Dendogram: `r params$show_row_dend`

# Paragraph

The following plots were generated using IDEP, including only the top `r params$n_genes` genes, using `r params$select_gene_id` to label, `r switch(params$dist_function, "1" = "Hierarchical", "2" = "K-Means")` distance,`r params$hclust_function` linkage, Cut off Z-score: `r params$heatmap_cutoff`, Center Genes: `r params$gene_centering`. genes are `r if(!params$gene_normalize) paste0(" not ")` normalized.

# Elbow Plot

```{r}
heatmap_data <- process_heatmap_data(
  data = params$pre_processed_data,
  n_genes_max = params$n_genes,
  #n_genes_min = 50,
  gene_centering = params$gene_centering,
  gene_normalize = params$gene_normalize,
  sample_centering = TRUE,
  sample_normalize = TRUE,
  all_gene_names = params$all_gene_names,
  select_gene_id = params$select_gene_id
)

k_means_elbow(heatmap_data = heatmap_data)
```

# Heatmap

```{r}
ht <- heatmap_main(
  data = heatmap_data,
  cluster_meth = params$cluster_meth,
  heatmap_cutoff = params$heatmap_cutoff,
  sample_info = params$sample_info,
  select_factors_heatmap = params$list_factors_heatmap,
  dist_funs = dist_functions(),
  dist_function = params$dist_function,
  hclust_function = params$hclust_function,
  no_sample_clustering = params$no_sample_clustering,
  heatmap_color_select = params$heatmap_color_select,
  row_dend = params$show_row_dend,
  k_clusters = params$k_clusters,
  re_run = FALSE
)
```

# SD Density

```{r sd_density}
sd_density(
  data = params$pre_processed_data,
  n_genes_max = params$n_genes
)
```

# Tree

```{r}
draw_sample_tree(
  tree_data = params$pre_processed_data,
  gene_centering = params$gene_centering,
  gene_normalize = params$gene_normalize,
  sample_centering = FALSE,
  sample_normalize = FALSE,
  hclust_funs = hcluster_functions(),
  hclust_function = params$hclust_function,
  dist_funs = dist_functions(),
  dist_function = params$dist_function
)
```

# Correlation Matrix

```{r}
cor_plot(
  data = params$pre_processed_data,
  label_pcc = TRUE,
  heat_cols = c("red", "black", "green"),
  text_col = "white"
)
```
